---
title: "Extract OSM features using R"
author: "Chia-Jung (Robert), Yeh"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: TRUE
    toc_depth: 3
    toc_float:
      collapsed: TRUE
      smooth_scroll: FALSE
    theme: lumen
---

<style type="text/css">

body{
font-size: 16px;
font-family:"Georgia","標楷體",Serif;
}
td {  /* Table  */
  font-size: 14px;
  font-family:Georgia,"標楷體",Serif;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
  font-family:Georgia,"標楷體",Serif;
}
h1 { /* Header 1 */
  font-size: 24px;
  font-family:Georgia,"標楷體",Serif;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family:Georgia,"標楷體",Serif;
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family:Georgia,"標楷體",Serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
.blogbody{
font-size:17px;
  font-family:Georgia,"標楷體",Serif;
line-height: 150%;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(kableExtra)
library(osmdata)
library(tmap)
library(sf)
tmap_mode("view")
```


# **Introduction**

## **What is OSM?**
OpenStreetMap (OSM) is an open-source platform that allows users to collaboratively edit and download geographic features with a standardised tagging structure. Each feature in OSM is described using a set of key-value tags. The **key** represents a general category of geographic elements, such as `amenity`, `highway`, `water.` The **value** represents a more specific classification within that category (key). For example, the key `highway` may include values like `motorway`, `trunk`, `primary`, and other road types. Therefore, this tagging system enables efficient structuring and extraction of geospatial features for elements across the globe.  

See [here](https://wiki.openstreetmap.org/wiki/Map_features) for all the tags for key and value in OSM.  

## **R package required**
The package required for connecting to OSM is `osmdata`. Please install the package as follows and load it into R code.  

```{r package, echo=T, eval=F}
install.packages("osmdata")
library(osmdata)
```

# **Using `osmdata`**

## **Basic structure**
The basic structure accessing OSM data through code is as follows. The `BOUNDARY_BOX` defines the spatial extent of the area from which we want to retrieve OSM data. Within the `add_osm_feature()` function, we need to specify the argument `key` and `value` to download the desired category of geographic features. Finally, using `osmdata_sf()` function to retrieve all the simple features (spatial data).  

```{r basic-code, echo=T, eval=F}
opq(BOUNDARY_BOX)%>%
  add_osm_feature(key="KEY", value="VALUE")%>%
  osmdata_sf()
```

The final results include `points`, `lines`, `polygons`, and also the multi-geometry.  


## **Example**

Now we take accessing the footpath data in Canberra for example.  

First of all, read the polygon data to define our target area. And please make sure that the CRS is EPSG::4326, which aligns with the CRS for OSM. If the CRS is not consistent, please use function `st_transform()` to convert it.     

```{r example1, echo=T, eval=T}
canberra=read_sf("https://raw.githubusercontent.com/tonywr71/GeoJson-Data/refs/heads/master/suburb-2-act.geojson")

print(st_crs(canberra)) # EPSG:4283
canberra=st_transform(canberra, crs=4326)
```

Then retrieve the boundary for the area using function `st_bbox()`.  

```{r example2, echo=T, eval=T}
# Extract the boundary of the spatial data
canberra_box=st_bbox(canberra)

print(canberra_box)
```

Let's plot a map to show the target area and its extent.  

```{r example3, echo=T, eval=T}
# Plot the map
tm_shape(canberra)+
  tm_polygons()+
  tm_shape(st_as_sfc(canberra_box))+
  tm_borders(lty="dashed", col="red")
```

We can then access OSM data using the basic structure mentioned above.   

```{r example4, echo=T, eval=T}
# Take sidewalk in Canberra for example
sidewalk=opq(canberra_box)%>%
  add_osm_feature(key="footway", value="sidewalk")%>%
  osmdata_sf()
```


Finally, plot all the spatial data to see the results.  

```{r example5, echo=T, eval=T}
# Plot the result
tm_shape(canberra)+
  tm_polygons(alpha=0.2, border.alpha=0.2)+
  tm_shape(sidewalk$osm_lines)+
  tm_lines(col="blue")+
  tm_shape(st_as_sfc(canberra_box))+
  tm_borders(lty="dashed", col="red")
```


