---
title: "R Preliminary"
author: "Chia Jung, Yeh\\\n National Yang Ming Chiao Tung University (NYCU), Taiwan"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: TRUE
    toc_depth: 3
    toc_float:
      collapsed: TRUE
      smooth_scroll: FALSE
    theme: lumen
---

<style type="text/css">

body{
font-size: 16px;
font-family:"Georgia","標楷體",Serif;
}
td {  /* Table  */
  font-size: 14px;
  font-family:Georgia,"標楷體",Serif;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
  font-family:Georgia,"標楷體",Serif;
}
h1 { /* Header 1 */
  font-size: 24px;
  font-family:Georgia,"標楷體",Serif;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family:Georgia,"標楷體",Serif;
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family:Georgia,"標楷體",Serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
.blogbody{
font-size:17px;
  font-family:Georgia,"標楷體",Serif;
line-height: 150%;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(kableExtra)
library(DT)
```


<p style="color:#003D79;font-size:16px">**[This notebook would introduce the basic syntax of R Programming.]**</p>


# **Object**

Object is the type of data. Commonly used object includes:  

1.  vector  
2.  factor
3.  matrix   
4.  data frame
5.  list 


## **Vector**
<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Features of vector**</p>  

* a combination of values
* one-dimensional
* it can be numeric, character or logical

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Use `c()` to create a vector**</p>  

* **Numeric vector**
```{r vector1, echo=T, eval=F}
vec1=c(1,2,3,4,5)
```

```{r vector1-show, echo=F, eval=T}
vec1=c(1,2,3,4,5)
vec1
```

* **Character vector**
```{r vector2, echo=T, eval=F}
vec2=c("A","B","C","D","E")
```

```{r vector2-show, echo=F, eval=T}
vec2=c("A","B","C","D","E")
vec2
```

* **Logical vector**
```{r vector3, echo=T, eval=F}
vec3=c(T,F,T,F,T)
```

```{r vector3-show, echo=F, eval=T}
vec3=c(T,F,T,F,T)
vec3
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Retrieve the specific elements**</p>  

* **Retrieve the third element in `vec2`**
```{r vector-retrieve1, echo=T, eval=T}
vec2[3]
```

* **Retrieve multiple elements in `vec2`**  
continuous elements  
```{r vector-retrieve2, echo=T, eval=T}
vec2[2:4]
```

* **Retrieve multiple elements in `vec2`**  
non-continuous elements
```{r vector-retrieve3, echo=T, eval=T}
vec2[c(1,3,5)]
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Create sequential numeric vector**</p> 

```{r vector-continuous1, echo=T, eval=T}
seq(from=2, to=20, by=2)
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Operations on vectors**</p>  

Let us set a numeric vector first.  
```{r vector-exg, echo=T, eval=T}
vec4=c(1,9,3,15,6,2,8,12)
```

* **Maximum & Minimum & Range**  
```{r vector-max-min, echo=T, eval=T}
max(vec4)
min(vec4)
range(vec4)
```

* **Sum & Mean & Median & Product**  
```{r vector-sum-mean, echo=T, eval=T}
sum(vec4)
mean(vec4)
median(vec4)
prod(vec4)
```

* **Variance & Standard deviation**  
```{r vector-std-var, echo=T, eval=T}
var(vec4)
sd(vec4)
```

* **Length**  
number of elements in the vector  
```{r vector-length, echo=T, eval=T}
length(vec4)
```

* **Order & Sort**  
sort would rearrange the sequence by ascending  
order would retrieve the index of ascending sequence  
```{r vector-sort-order, echo=T, eval=T}
sort(vec4)
order(vec4)
```

See the result of `order()` above. It shows that the last value is "4", which means that the index of largest value is "4", that is, `vec4[4]` (15).  


* **Unique**  
remove the duplicated element   
```{r vector-unique, echo=T, eval=T}
vec_dup=c(1,9,5,2,6,1,8,5,2)
unique(vec_dup)
```

* **Check NA**  
check if the element is NA  
```{r vector-na, echo=T, eval=T}
vec_na=c(1,9,5,NA,6,NA)
is.na(vec_na)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Calculations with vectors**</p>  

Let us use `vec4` and set a new numeric vector `vec5`.  
```{r vector-vec5, echo=T, eval=T}
vec5=c(2,5,8,11,7,4,10,3)
```

* **Arithmetic (both vectors)**    
add/ multiple two vectors  
<p style="color:#CE0000;font-size:14px;line-height:1">**Note that the lengths of the vectors should be the same while doing the calculations!**</p>  
```{r vector-addition1, echo=T, eval=T}
vec4+vec5
vec4*vec5
```

* **Arithmetic (vector & element)**  
add/ multiple a value to all the elements in the vector  
```{r vector-addition2, echo=T, eval=T}
vec5+5
vec5*5
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Convert data type**</p>  
* **From character to numeric**  
<p style="color:#CE0000;font-size:14px;line-height:1">**Note that sometimes we may get a "character" type of number when importing the data. We should convert them into "numeric" type in advance, so as to do calculations.**</p>  
```{r vector-convert1, echo=T, eval=T}
vec_cha=c("1","2","3","4","5")
```

* **Use `class()` to check out the data type**  

```{r vector-convert2, echo=T, eval=T}
class(vec_cha)
```

* **Convert it by function `as.numeric()`**  
```{r vector-convert3, echo=T, eval=T}
as.numeric(vec_cha)
```

Also, we can convert the numeric to character as well. Use the function `as.character()`.  



## **Factor**
<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Features of factor**</p>  

* converted from the character vector  
* should set the levels  
* some useful tips in R:  
  * setting a factor is very useful when adjusting the sequence of legend in `ggplot2`  
  * used to set the dummy variable in the econometric model  

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Use `factor()` to create a factor**</p>  

* **Function & Argument**
```{r factor1, echo=T, eval=F}
factor(c(...), levels=c(...))
```

Let us set a vector of the schools:  
```{r factor2, echo=T, eval=T}
school=c("University","Kindergarten","Elementary","High","Junior High")
```

We know that the sequence should be Kindergarten >> Elementary >> Junior High >> High >> University.

* **Use `factor()` to set their levels**  
```{r factor3, echo=T, eval=T}
school=factor(school, levels=c("Kindergarten","Elementary","Junior High","High","University"))
```

```{r factor4, echo=F, eval=T}
school
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Check if `school` is the factor**</p>  

```{r factor5, echo=T, eval=T}
is.factor(school)
class(school)
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Calculate number of elements in each level**</p>  

```{r factor6, echo=T, eval=T}
summary(school)
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Convert factor to numeric**</p>  

* **Use `as.numeric()` to transform the factor to numeric**    
```{r factor7, echo=T, eval=T}
as.numeric(school)
```

The numeric retrieved above represents the level of each elements.  


## **Matrix**
<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Features of matrix**</p>  

* contains multiple rows and columns 
* combination of multiple vectors with the same types (numeric, character or logical)  

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Use `matrix()` to create a matrix**</p>  

* **Function & Argument**
  * `nrow=` means number of rows in the matrix  
  * `ncol=` means number of columns in the matrix  
```{r matrix-function, echo=T, eval=F}
matrix(c(...), nrow=, ncol=)
```

```{r matrix1, echo=T, eval=F}
mat1=matrix(c(1:15), nrow=3, ncol=5)
```

```{r matrix1-show, echo=F, eval=T}
mat1=matrix(c(1:15), nrow=3, ncol=5)
mat1
```

See the result above, the original vector (`c(1:15)`) is filled by columns. What if we want it by row?

* **Set argument `byrow=T`**

```{r matrix2, echo=T, eval=F}
mat2=matrix(c(1:15), nrow=3, ncol=5, byrow=T)
```

```{r matrix2-show, echo=F, eval=T}
mat2=matrix(c(1:15), nrow=3, ncol=5, byrow=T)
mat2
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Check out the number of rows, columns and dimension of matrix**</p>  

```{r matrix3, echo=T, eval=T}
nrow(mat2)
ncol(mat2)
dim(mat2)
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Retrieve the specific elements in the matrix**</p>  

The index of row should be set at first, and then set the index of column.  
```{r matrix4, echo=T, eval=T}
mat2[2, 4]
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Transpose the matrix**</p>  

```{r matrix5, echo=T, eval=T}
t(mat2)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Calculations on matrix**</p>  

* **Sum of each row/ column**
```{r matrix6, echo=T, eval=T}
rowSums(mat2)
colSums(mat2)
```

* **Mean of each row/ column**  
```{r matrix7, echo=T, eval=T}
rowMeans(mat2)
colMeans(mat2)
```


## **Data frame**

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Features of data frame**</p>  

* like a matrix but can have columns with different types  
* works like a spreadsheet (Excel)
* commonly used in data analysis
* can be tidy up by the package `dplyr` (see [Section 2](#tidy-data))


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Use `data.frame()` to create a data frame**</p>  

* **Function & Argument**

```{r dataframe-function, echo=T, eval=F}
data.frame(A=c(), B=c(), C=c())
```

`A`, `B`, and `C` in the function represent the column names of the table.  


```{r dataframe1, echo=T, eval=T}
StuScore=data.frame(StudentID=c("ID1","ID2","ID3","ID4","ID5"),
           Name=c("Bob","Mary","Robert","Jason","Jane"),
           Score=c(60,80,40,50,100))
```

```{r dataframe1-show, echo=F, eval=T}
StuScore
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Check out the column names of data frame**</p>  
```{r dataframe2, echo=T, eval=T}
colnames(StuScore)
```

<a name="dataframe-rowcol">
<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Retrieve specific rows/ columns**</p>  
</a>

* Retrieve rows  
place the index in front of the comma
```{r dataframe3, echo=T, eval=T}
StuScore[c(1,3,5),]
```

* **Retrieve columns**  
place the index after the comma
```{r dataframe4, echo=T, eval=T}
StuScore[, c(1,2)]
```

* **Retrieve columns by specifying the column names**  
place the column name vector after the comma
```{r dataframe4-1, echo=T, eval=T}
StuScore[, c("StudentID","Name")]
```

* **Retrieve single column by `data$column_name`**  
```{r dataframe5, echo=T, eval=T}
StuScore$Name
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Retrieve specific element**</p>  

* **Set the index of row and column**  
`data[row_index, column_index]`
```{r dataframe6, echo=T, eval=T}
StuScore[2,2]
```

* **Edit the element by giving value**  
```{r dataframe7, echo=T, eval=T}
StuScore[2,2]="Jessica"

# check out the new data
StuScore
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Check out the first/ last 6 rows of data**</p>  

Here we use the data `iris` provided by the R base. `iris` is a well-known data set that consists of 150 samples from each of three species of Iris flower (setosa, virginica, and versicolor). The columms include the length and the width of the sepals and petals (5 columns in total).  

* **Use `head()` to retrieve the first 6 rows**  
```{r dataframe_head, echo=T, eval=T}
head(iris)
```

* **Use `tail()` to retrieve the last 6 rows**  
```{r dataframe_tail, echo=T, eval=T}
tail(iris)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Extend the data frame**</p>  

Back to our `StuScore` example. If we want to add attributes of a new student (add new row):

* **Use `rbind()` to add new rows**  
```{r dataframe_rbind, echo=T, eval=T}
# first create a new data
new_student=data.frame(StudentID="ID6", Name="Bob", Score=90)

# use rbind to merge data
StuScore=rbind(StuScore, new_student)

# check out the new data
StuScore
```

Then, if we want to add a new attribute (e.g., Gender) for all students (add new column):

* **Use `cbind()` to add new columns**  
```{r dataframe_cbind, echo=T, eval=T}
# first create a new data
Gender=c("M","F","M","M","F","M")

# use cbind to merge data
StuScore=cbind(StuScore, Gender)

# check out the new data
StuScore
```

Or it can simply add by using: `data$new_column=c(...)`
```{r dataframe_add_col, echo=T, eval=T}
StuScore$Height=c(180,165,170,160,175,180)

# check out the new data
StuScore
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Check out the duplicated rows**</p>  

Let's create a new data with duplicated rows.  

```{r dataframe_duplicated1, echo=T, eval=T}
StuScore_dup=data.frame(StudentID=c("ID1","ID2","ID4","ID3","ID4","ID5","ID2"),
           Name=c("Bob","Mary","Jason","Robert","Jason","Jane","Mary"),
           Score=c(60,80,40,100,40,100,80))
```

```{r dataframe_duplicated1-show, echo=F, eval=T}
StuScore_dup
```

We can see some of data are duplicated (ID2 & ID4). Let's use `duplicated()` to check out whether it is repeated for each row.  
```{r dataframe_duplicated2, echo=T, eval=T}
duplicated(StuScore_dup)
```

The result shows that the row 5 and 7 are duplicated (`TRUE`). Note that the first appeared row would not be viewed as duplicated.  



## **List**

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Features of list**</p>  

* collection of objects, which can be vector, matrix, or data frame  
* most complicated, many packages use list to return the result  


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Use `list()` to create a list**</p>  

```{r list1, echo=T, eval=T}
StuScore_list=list(StudentID=c("ID1","ID2","ID3","ID4","ID5"),
           Name=c("Bob","Mary","Robert","Jason","Jane"),
           Score=c(60,80,40,50,100),
           Class="A")
```

```{r list1-show, echo=F, eval=T}
StuScore_list
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾  Retrieve the attribute of list**</p>  

* **Use `list$NAME` to retrieve**  
```{r list2, echo=T, eval=T}
StuScore_list$Score
```

* **Use index to retrieve**  
```{r list3, echo=T, eval=T}
StuScore_list[[2]]
```



# **Tidy data**
The package commonly used to tidy data frame is `dplyr`. Please install and import the package via the following code:  

```{r install-dplyr, echo=T, eval=F}
# install the package
install.packages("dplyr")

# import the package
library(dplyr)
```

Functions that would be introduced in this section are summarized in Table \@ref(tab:function-dplyr).  

```{r function-dplyr, echo=F, eval=T}
fuc=data.frame(Functions=c("`select()`","`filter()`","`mutate()`","`group_by()`","`summarise()`","`left_join()`","`arrange()`","`rename()`","`distinct()`"), Usage=c("select the specific column of data frame", "filter the data frame by given conditions", "add a new column (attribute) in data frame", "group the data by specific columns, it is commonly followed by `filter()` or `summarise()`", "conduct the basic calculation on data frame, it is commonly used with `group_by()`", "join two data by a specific key (column)", "arrange the data by specific columns", "rename the column name of data frame", "remove the duplicated rows"))

kbl(fuc, booktabs=T, escape=F, caption="Vital Functions in `dplyr`")%>%
  kable_styling(bootstrap_options=c("striped", "hover"), font_size=14)
```

Before introducing how to process the data by using package `dplyr`, let's first see how to read the data.  

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Read data by `read.csv()`**</p>  

* **From local directory**  
when the file is in your computer  
```{r read-local, echo=T, eval=F}
# function
read.csv("DIRECTORY")

# example
read.csv("C:/Users/Robert/Downloads/world.csv")
```

<p style="color:#CE0000;font-size:14px;line-height:1">**Note that the "single forward slash (/)" is used in the directory.<br>Do not use "single backslash (\\)!<br>Alternatively, we can use "double backslash (\\\\)" instead.**</p>  

* **From website**  
when the file is on the website (with a url)  
```{r read-url, echo=T, eval=F}
# function
read.csv("url")

# example
read.csv("https://raw.githubusercontent.com/ChiaJung-Yeh/R-Code-Sharing/main/R_Preliminary/world.csv")
```


Here we adopt the second method to obtain our data.  
```{r read-data, echo=T, eval=F}
# read data
world=read.csv("https://raw.githubusercontent.com/ChiaJung-Yeh/R-Code-Sharing/main/R_Preliminary/world.csv")

# check out the data
world
```


```{r read-data-show, echo=F, eval=T, message=F, warning=F}
world=read.csv("https://raw.githubusercontent.com/ChiaJung-Yeh/R-Code-Sharing/main/R_Preliminary/world.csv")
datatable(world, options=list(pageLength=5, scrollX=T, headerCallback = DT::JS(
    "function(thead) {",
    "  $(thead).css('font-size', '0.7em');",
    "}"
  )))%>%
  formatStyle(columns=c(1:ncol(world)), fontSize='12px')%>%
  formatRound(c("area_km2", "lifeExp","gdpPercap"), digits=2)
```


## **Select**
Select the column we want in the data frame. This function is the same as retrieving columns in data frame introduced in [Section 1.4](#dataframe-rowcol).  

* **Function & Argument**

```{r select-function, echo=T, eval=F}
select(DATA, col1, col2, ...)
```

`col1` and `col2` above represent the column names that would be retrieved.  

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Select columns by `select()`**</p>  

```{r select1, echo=T, eval=T}
# select data
world_sel1=select(world, name_long, continent)

# check out the first 6 data
head(world_sel1)
```

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Select columns by given column names vector**</p>  

First set up a character vector to store the column name that would be retrieved. Then, Use `all_of()` in `select()` to call all of the column stored in the vector.  

```{r select2, echo=T, eval=T}
# specified columns vector
sp_col=c("iso_a2", "region_un")

# select data with specified columns
world_sel2=select(world, all_of(sp_col))

# check out the first 6 data
head(world_sel2)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Delete columns in data frame**</p>  

To do so, just simply add a minus (`-`) before the column name that should be removed.  

```{r select3, echo=T, eval=T}
# delete column
world_sel3=select(world, -continent, -region_un, -subregion, -type)

# check out the first 6 data
head(world_sel3)
```


## **Filter**
Filter the data frame by given conditions.  

* **Function & Argument**

```{r filter-function, echo=T, eval=F}
filter(DATA, condition1, condition2, ...)
```

`condition1` and `condition2` above represent the condition we set. Add more conditions in `filter()` means that **ALL** the conditions should be satisfied (condition1 & condition2 & ...).  


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Numeric filter**</p>  

filter the country where its population (`pop`) is over 100 million
```{r filter1, echo=T, eval=T}
# filter data by population
world_fil1=filter(world, pop>100000000)

# check out the first 6 data
head(world_fil1)

# check how many countries satisfy with the requirement
nrow(world_fil1)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Character filter**</p>  
`A %in% B` is used to check if each of the element in vector A is listed in the vector B  
  
filter the country where its continent (`continent`) is in Asia or Europe  
```{r filter2, echo=T, eval=T}
# filter data by population
world_fil2=filter(world, continent %in% c("Asia", "Europe"))

# check out the first 6 data
head(world_fil2)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Multiple filter (AND)**</p>  
* add conditions in the function `filter()`  
* alternatively, we can use `&` to connect all of the conditions in `filter()`  
  
filter the country where its continent (`continent`) is in Asia or Europe **AND** the population (`pop`) is over 100 million
```{r filter3, echo=T, eval=T}
# filter data by multiple conditions
world_fil3=filter(world, continent %in% c("Asia", "Europe"), pop>100000000)

# check out the first 6 data
head(world_fil3)

# check how many countries satisfy with the requirements
nrow(world_fil3)
```

Or, we can use `&` to connect the conditions as follows.  
```{r filter4, echo=T, eval=T}
# filter data by multiple conditions
world_fil4=filter(world, continent %in% c("Asia", "Europe") & pop>100000000)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Multiple filter (OR)**</p>  
* when any of the condition is met, retrieve the data  
* use `|` to connect all of the conditions in `filter()`  
```{r filter5, echo=T, eval=T}
# filter data by multiple conditions
world_fil5=filter(world, continent %in% c("Asia", "Europe") | pop>100000000)

# check how many countries satisfy with the requirements
nrow(world_fil5)
```



## **Mutate**
Add the new columns (attributes) or update the columns in data frame. This function is similar to `cbind()` introduced in [Section 1.4](#dataframe-rowcol).  

* **Function & Argument**

```{r mutate-function, echo=T, eval=F}
mutate(DATA, A=..., B=..., ...)
```

`A` and `B` above represent the operation of generating the new columns.  

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Mutate a new column by column operations**</p>  

add a column to calculate the population density of each country (`pop_dens`)  
<p style="color:#CE0000;font-size:14px;line-height:1">**Population density is population (`pop`) divided by area (`area_km2`).**</p>  

```{r mutate1, echo=T, eval=T}
# mutate a column for population density
world_mut1=mutate(world, pop_dens=pop/area_km2)

# check out the first 6 data
head(world_mut1[, c("name_long","continent","pop","area_km2","pop_dens")])
```

<p style="color:#CE0000;font-size:14px;line-height:1">**Since there are missing data on population in some of countries, there exists NA in `pop_dens`.**</p>  


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Mutate a new column with `ifelse()`**</p>  

* **Function & Argument**

```{r ifelse-function, echo=T, eval=F}
ifelse(condition, do when true, do when false)
```


add a column to label the country if area and population are both higher than median (label 1 if true)  
<p style="color:#CE0000;font-size:14px;line-height:1">**Note that there are NA in the column, we should set `na.rm=T` in the function `median()`, to remove NA in advance.**</p>  

```{r mutate2, echo=T, eval=T}
# calculate the medium first
pop_med=median(world$pop, na.rm=T)
area_med=median(world$area_km2, na.rm=T)

# mutate a column for label
world_mut2=mutate(world, pop_area=ifelse(pop>pop_med & area_km2>area_med, 1,0))

# check out the first 6 data
head(world_mut2[, c("name_long","continent","pop","area_km2","pop_area")])
```



<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Mutate a new column with `case_when()`**</p>  

* **Function & Argument**

```{r case_when-function, echo=T, eval=F}
case_when(DATA, 
          condition1 ~ result1, 
          condition2 ~ result2, 
          condition3 ~ result3,...)
```

It means that if the `condition` is satisfied, then retrieve the `result`.  


add a column to classify the country into large area large population (LALP), large area small population (LASP), small area large population (SALP), and small area small population (SASP) by the median of each attribute  

```{r mutate3, echo=T, eval=T}
# calculate the medium first
pop_med=median(world$pop, na.rm=T)
area_med=median(world$area_km2, na.rm=T)

# mutate a column for classfication
world_mut3=mutate(world, pop_area_cla=case_when(
  pop>pop_med & area_km2>area_med ~ "LALP",
  pop>pop_med & area_km2<area_med ~ "LASP",
  pop<pop_med & area_km2>area_med ~ "SALP",
  pop<pop_med & area_km2<area_med ~ "SASP"))

# check out the first 6 data
head(world_mut3[, c("name_long","continent","pop","area_km2","pop_area_cla")])
```


## **Group by**

`group_by()` is usually followed by `filter()` or `summarise()`. To connect two different function, package `dplyr` provides a useful tools, pipe (`%>%`), to connect the functions. When using it, data should only place in the first function.  

* **Function & Argument**  

`group_by() %>% filter()`  
```{r groupby-function1, echo=T, eval=F}
group_by(DATA, column)%>%
  filter(condition)
```

`group_by() %>% summrise()`  
```{r groupby-function2, echo=T, eval=F}
group_by(DATA, column)%>%
  summarise(operation)
```


<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Group by with filter**</p>  

filter the largest population (`pop`) of each continent  

```{r groupby1, echo=T, eval=T, warning=F}
# calculate the medium first
world_gro1=group_by(world, continent)%>%
  filter(pop==max(pop, na.rm=T))

# check out the result
world_gro1
```

By using `group_by() %>% filter()`, all of the attributes would be retained.  

<p style="color:#003D79;font-size:18px;line-height:2">**⌾ Group by with summarise**</p>  
calculate the total population (`pop`) and average life expectancy (`lifeExp`) of each continent  

```{r groupby2, echo=T, eval=T, warning=F}
# calculate the medium first
world_gro2=group_by(world, continent)%>%
  summarise(total_pop=sum(pop, na.rm=T),
            ave_lifeExp=mean(lifeExp, na.rm=T))

# check out the result
world_gro2
```

By using `group_by() %>% summarise()`, it can retrieve the operation on specific column. Note that only the specified column are retained in the result.  


## **Left join**
Merge two different data by referencing the key (terminology in database management: primary key)

* **Function & Argument**  

```{r leftjoin-function, echo=T, eval=F}
left_join(DATA1, DATA2, by=c("col1"="col2"))
```

It means that using `col1` in `DATA1` and `col2` in `DATA2` to merge two data.  

Here we use a 






