---
title: "ggplot2繪圖應用"
author: "Chia-Jung, Yeh"
date: "`r Sys.Date()`"
output: html_document
---

<style type="text/css">

body{
font-size: 18px;
font-family:"Times New Roman","標楷體",Georgia,Serif;
}
td {  /* Table  */
  font-size: 8px;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
}
h1 { /* Header 1 */
  font-size: 28px;
      font-family:"Times New Roman","標楷體",Georgia,Serif;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family:"Times New Roman","標楷體",Georgia,Serif;
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
.blogbody{
font-size:17px;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
line-height: 150%;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```


# **ggplot2**

```{r ggplot2, echo=T, eval=T, error=T, fig.width=6, fig.height=8, dpi=100, fig.align="center"}

```





