---
title: "OSM Restaurant"
output: html_document
---

<style type="text/css">

body{
font-size: 18px;
font-family:"Times New Roman","標楷體",Georgia,Serif;
}
td {  /* Table  */
  font-size: 8px;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
}
h1 { /* Header 1 */
  font-size: 28px;
      font-family:"Times New Roman","標楷體",Georgia,Serif;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family:"Times New Roman","標楷體",Georgia,Serif;
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
.blogbody{
font-size:17px;
  font-family:"Times New Roman","標楷體",Georgia,Serif;
line-height: 150%;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(tmap)
library(sf)
library(dplyr)

tmap_mode("view")

```

```{r map, eval=T, echo=F, out.width='100%'}
amenity=fread("D:/Sidewalk/Analysis/5_POI Categorization/C_OpenStreetMap/amenity.csv")%>%
  mutate(geometry=st_as_sfc(paste0("POINT(", lon, " ", lat, ")")))%>%
  st_sf(crs=4326)%>%
  st_transform(crs=3826)

tm_shape(filter(amenity, amenity=="restaurant"))+
  tm_dots()

```

